Boccaccio example 2: end to end truth for California Current

### 2019-06-14 FullBoccaccioTruthEx2.Rmd
try bocaccio new CCA age splitting

---
title: "Boccaccio example: end to end truth for California Current"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

  * "CalCurrent2013_OA_off": California Current Atlantis model output from Isaac Kaplan's google drive \outputFolderAllOff2013Oceanography using the Hodgson et al. 2018 configuration. Corresponds to CC1Config.R file.

@@

@@

# call sample_survey_biomass with a bunch of 1s for weight at age
# in the code it multiplies atoutput by wtatage so this allows us to use
# biomass directly
wtage <- data.frame(species=rep(survspp, each=10),
                    agecl=rep(c(1:10),length(survspp)),
                    wtAtAge=rep(1.0,length(survspp)*10))

@@

surveyB_frombio_ss <- surveyB_frombio[surveyB_frombio$species == "Bocaccio_rockfish",]

@@

atBtxt2tidy <- atBtxt2 %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% c("Bocaccio_rockfish"))

#all species comparison, time intervals hardcoded for 5 steps per year
compareB <-ggplot() +
  geom_line(data=surveyB_frombio_ss, aes(x=time/5,y=atoutput, color="survey census B"), 
            alpha = 10/10) +
  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 1/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

@@

surveyN_ss <- surveyN[surveyN$species == "Bocaccio_rockfish",]

plotN <-ggplot() +
  geom_line(data=surveyN_ss, aes(x=time/5,y=atoutput, color="survey census N"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

@@

```{r truenatagecl, eval=FALSE}
# We need an effective N for the sample fish function, setting equal to actual N
# this one is high but not equal to total for numerous groups
effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))

numsallhigh <- sample_fish(survey_testNall, effNhigh)

@@

```{r testagecl, eval = FALSE}
# add YOY file to the config files
YOY <- load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))

@@

@@

# test with one species--this is still full polygon and layer output
trueNatage_ss <- trueNatage[trueNatage$species == "Bocaccio_rockfish",]

@@

len <- length_census_ss$natlength %>%
  filter(species == "Bocaccio_rockfish")

@@

truecatchbio_ss <- truecatchbio[truecatchbio$species == "Bocaccio_rockfish",]

@@

Now (that catch nums are corrected!) we can sample fishery catch for lengths, ages, weights:

@@

Now we know the fishery output is annual (365 days), and there will only be 100 years. Because we don't want to wait 24 hours for this, we will look at only our three species.

@@

This is true catch at age for bocaccio:

@@

>>>>

---
title: "Boccaccio example 2: end to end truth for California Current"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

  * "CalCurrent2063_run11": California Current Atlantis model output from Isaac Kaplan's google drive \output_CC_2063_run11 using the Hodgson et al. 2018 configuration, and driven with the F time series and recruitment variability we plan to use for our Fragile Ecosystems, Robust Assessments project. Corresponds to CC2Config.R file.

@@

species_ss <- "Bocaccio_rockfish"

@@

# call sample_survey_biomass with a bunch of 1000s for weight at age
# in the code it multiplies atoutput by wtatage so this allows us to use
# biomass directly
wtage <- data.frame(species=rep(survspp, each=10),
                    agecl=rep(c(1:10),length(survspp)),
                    wtAtAge=rep(1000.0,length(survspp)*10))

@@

surveyB_frombio_ss <- surveyB_frombio[surveyB_frombio$species == species_ss,]

@@

atBtxt2tidy <- atBtxt2 %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species %in% species_ss)

#all species comparison, time intervals hardcoded for 5 steps per year
compareB <-ggplot() +
  geom_line(data=surveyB_frombio_ss, aes(x=time/stepperyr,y=atoutput, color="survey census B"), 
            alpha = 10/10) +
  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 1/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

@@

surveyN_ss <- surveyN[surveyN$species == species_ss,]

plotN <-ggplot() +
  geom_line(data=surveyN_ss, aes(x=time/stepperyr,y=atoutput, color="survey census N"), 
            alpha = 10/10) +
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

@@

```{r truenatagecl, eval=FALSE}
# We need an effective N for the sample fish function, setting equal to actual N
# this one is high but not equal to total for numerous groups
survey_testNall <- readRDS(file.path(d.name, paste0(scenario.name, "survey_testNall.rds")))

numsall <- aggregate(survey_testNall$atoutput,list(survey_testNall$species,survey_testNall$agecl,survey_testNall$time),sum)
names(numsall) <- c("species","agecl","time","atoutput")
numsallhigh <- data.frame(species = numsall$species,
					  agecl = numsall$agecl,
					  polygon = NA,
					  layer = NA,
					  time = numsall$time,
					  atoutput = numsall$atoutput)

# better to do an actual census above, can't make effNhigh high enough for some spp
#effNhigh <- data.frame(species=survspp, effN=rep(1e+8, length(survspp)))
#numsallhigh <- sample_fish(survey_testNall, effNhigh)

@@

```{r testagecl, eval=FALSE}
# add YOY file to the config files
YOY <- load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))

@@

Calc Z is breaking for one of the many species; Isaac says something like tuna, sharks, or salmon may not fit this design, so let's check who is having problems:

```{r calcZsnippetagain, eval=FALSE}

yoy <- YOY
nums <- survey_testNall
fgs <- funct.groups
biolprm <- biol
toutinc <- runpar$toutinc

#calc_Z <- function(yoy, nums, fgs, biolprm, toutinc) {
  # subset the yoy for species included in the fgs file
  # that are turned on
  # colnames of the recruit data are "Time" and a column for each
  # species where the name is the functional group with an additional .0
  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
  recruits <- yoy[, colnames(yoy) %in% c("Time", paste0(turnedon$Code, ".0"))]

  # mg carbon converted to wet weight in tonnes
  k_wetdry <- biolprm$kgw2d / 1000000000
  # Sum of structural and reserve nitrogen (KWSR_RN + KWSR_SN)
  nitro <- merge(biolprm$kswr, biolprm$kwrr, by = "1")
  nitro$sum <- apply(nitro[, 2:3], 1, sum)

  # legacy: If output is from legacy code there will be an error in the
  # yoy data, where the first row is in a different unit.
  # yoy.txt is the biomass in tonnes per spawning event summed over the total
  # model domain.
  # The first row (< Nov/Dec 2015) is stored as biomass and the remaining rows
  # are stored in numbers, must convert the entire matrix to biomass
  # Check if legacy code and if so convert the numbers to biomass

  # G.Fay 2/21/16 : changed yoy to recruits in below loop.
  if (abs(recruits[1, 2] / recruits[2, 2]) > 10) {
    recruits[2:NROW(recruits), 2:NCOL(recruits)] <- recruits[2:NROW(recruits), 2:NCOL(recruits)] *
    nitro[match(gsub(".0", "", colnames(recruits)[-1]), nitro[, 1]), "sum"]
  }

  # Wide to long
  recruits <- reshape(data = recruits, direction = "long",
    varying = colnames(recruits)[-1],
    v.names = "recruits",
    times = colnames(recruits)[-1],
    timevar = "group")
  rownames(recruits) <- 1:NROW(recruits)
  recruits <- recruits[, -which(colnames(recruits) == "id")]
  # Switch from species code to species
  recruits$group <- gsub("\\.0", "", recruits$group)
  recruits <- merge(recruits, fgs[, c("Code", "Name")],
    by.x = "group", by.y = "Code")

  # merge recruits with strucn and resn of recruits from biol.prm file
  recruits <- merge(recruits, nitro[, c(1, 4)],
    by.x = "group", by.y = "1", all.x = TRUE, all.y = FALSE)
  colnames(recruits)[which(colnames(recruits) == "recruits")] <- "recruitsbio"
  # Get recruits in numbers rather than biomass
  recruits$recruits <- recruits$recruitsbio / recruits$sum
  recruits$yr <- as.integer(round(recruits$Time)/365) #needed to merge with totnums

  # June 2019:Isaac determined that yr 0 (Time 0) in the YOY file is not real, don't use it
  # code below assigns year 1 to times 0:stepperyr, all good output timesteps in truth$nums

  # G.Fay 2/21/16
  # UGLY code below tries to align fraction of annual yoy with timing of recruitment
  # values in YOY.txt are total YOY that year waiting to recruit.
  # seems to get rid of most of 'issues' - some v.minor survival >1,
  #perhaps due to averaging of survival over toutinc days
  nyrs <- ceiling(max(yoy$Time)/365)
  times <- unique(yoy$Time)

  # SKG June 2019
  # need to match codes below, this mismatches when biolprm not in same order!!
  #recstart_temp <- biolprm$recruit_time
  #recstart_temp[,2] <- biolprm$time_spawn[-(grep('#',
  #                     biolprm$time_spawn[,1])),2] + biolprm$recruit_time[,2]
  # was this grep to get rid of a species with #XXX? breaks when there are none
  #recstart_temp[,2] <- biolprm$time_spawn[,2] + biolprm$recruit_time[,2]

  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%turnedon$Code,] #bug added codes with no YOY output
  #recstart_temp <- recstart_temp[recstart_temp[,1]%in%recruits$group,]

  # June 12 2019 Beth determined that we do not need spawn_period

  # rectiming is a dataframe with species code, time_spawn (day of year),
  # recruit_time (number of days), and recruit period (number of days).
  # we use these to calculate recstart (day of year) and recend (day of year)
  rectiming <- merge(biolprm$time_spawn, biolprm$recruit_time, by = 1)
  rectiming <- merge(rectiming, biolprm$recruit_period, by = 1)
  names(rectiming) <- c("Code", "time_spawn", "recruit_time", "recruit_period")
  rectiming <- rectiming %>%
    mutate(recstart = time_spawn + recruit_time) %>%
    mutate(recend = recstart + recruit_period)

  # Sum numbers output over all boxes/depth/cohorts
  # align model output timesteps (days) with recruitment periods (days)
  totnums <- aggregate(atoutput ~ species + time, data = nums, sum) %>%
    #mutate(time.days = (time+1)*toutinc) %>% #makes time 0 into days 0->73, etc
    mutate(time.days = (time)*toutinc) %>% #makes time 1 into days 0->73, etc
    mutate(yr = ceiling(time.days/365))  # yr 1 is 0:stepsperyr to match recruits yr1

  totnums <- merge(totnums, recruits,
                   by.x = c("yr", "species"), by.y = c("yr", "Name"),
                   all.x = TRUE) %>%
    arrange(time)

  #subset recpars to groups of interest
  recstart_temp <- rectiming[rectiming$Code %in% totnums$group,]

  totnums$frac_recruit <- 0

  for (irow in 1:nrow(recstart_temp)) {
    group <- recstart_temp$Code[irow]
    pick <- which(totnums$group == group)

    recstart <- seq(recstart_temp$recstart[irow],by=365,length.out=nyrs)
    recstart <- recstart[recstart<max(totnums$Time[pick])]
    recend <- recstart + recstart_temp$recruit_period[irow]
    #rec_times <- rbind(rec_times,cbind(group,recstart,recend))

    for (i_rec in 1:length(recstart)) {
      i_tstart <- which(pick==min(pick[totnums$time.days[pick]>=recstart[i_rec]]))
      i_tstop <- which(pick==min(pick[totnums$time.days[pick]>=recend[i_rec]]))
      if (i_rec == length(recstart)) i_tstop <- length(pick)
      n_t <- 1+i_tstop-i_tstart
      for (i_t in 1:n_t) {
        t_temp <- totnums$time.days[pick[i_tstart+i_t-1]]
        num_temp <- t_temp - recstart[i_rec]
        if (i_t>1) {
          if ((recend[i_rec]-t_temp)>toutinc) {
            num_temp <- toutinc
          }
          else {
            num_temp <- recend[i_rec]-(totnums$time.days[pick[i_tstart+i_t-2]])
          }
        }
        frac_temp <- max(c(0,num_temp /(recend[i_rec]-recstart[i_rec])))
        if(frac_temp > 1.0) frac_temp = 1.0
        totnums$frac_recruit[pick[i_tstart+i_t-1]] <- frac_temp
      }
    }
  }

  #G.Fay 1/6/16, expand to num of recruits
  # Recruit / mg C converted to wet weight in tonnes / redfield ratio of C:N
  totnums$recruits <- totnums$recruits/k_wetdry/biolprm$redfieldcn

  totnums$group <- recruits$group[match(totnums$species, recruits$Name)]
  # For all time increments where there
  totnums$recruits[is.na(totnums$recruits)] <- 0

  totnums$annrecruits <- totnums$recruits

  totnums$recruits <- totnums$annrecruits*totnums$frac_recruit

  # Calculate survivors for each species group
  totnums$survivors <- totnums$recruits

  totnums$survival <- totnums$survivors
  # Calculate survival for each group
  for (group in unique(totnums$group)) {
    #if(group == "SHD") browser()
    pick <- which(totnums$group == group)
    survival_temp <- c(NA,
      totnums$survivors[pick[-1]]/totnums$atoutput[pick[-length(pick)]])

    # G.Fay 2/21/16  "think" this is what things should be, recruits don't show up
    # in numbers at age until time step after the recruitment event.
     survival_temp <- c(
       (totnums$atoutput[pick[-1]]-
          totnums$recruits[pick[-1]])/totnums$atoutput[pick[-length(pick)]],NA)

#     survival_temp <- c(
#       (totnums$atoutput[pick[-1]])/
#          (totnums$recruits[pick[-1]]+totnums$atoutput[pick[-length(pick)]]),NA)

    survival_temp[survival_temp < 0] <- NA
    # Use first positive value to replace the initial year and all negative vals
    firstgood <- which(!is.na(survival_temp))[1]

    survival_temp[1:firstgood] <- survival_temp[firstgood]
    for(ii in seq_along(survival_temp)) {
      if (is.na(survival_temp[ii])) {
        nonzero <- which(which(survival_temp > 0) > ii)
        if (length(nonzero) == 0) nonzero <- which(survival_temp > 0)
        survival_temp[ii] <- survival_temp[which.min(abs(nonzero - ii))]
      }
    }
    totnums$survival[pick] <- survival_temp
   }

  #Calculate Z
  totnums$Z <- -1 * log(totnums$survival)
  finaldata <- data.frame("species" = totnums$species,
    "agecl" = NA, "polygon" = NA, "layer" = NA,
    "time" = totnums$time, "atoutput" = totnums$Z)

  #return(finaldata)
#}

Modified the `calc_Z` function so it only evaluates species in the input numbers file. This creates Z values for all species. However, the call to `calc_stage2age` returns only a few of the 36 species from calc_Z so lets check the problem here:

```{r calcstage2agesinppetagain, eval=FALSE}

yoy <- YOY
nums_data <- survey_testNall
fgs <- funct.groups
biolprm <- biol
runprm = runpar

#calc_stage2age <- function(nums_data, biolprm, yoy, fgs, runprm) {

  # subset the yoy for those species that are included in the fgs file
  # that are turned on
  species.code <- fgs$Code
  turnedon <- fgs[fgs$IsTurnedOn > 0, ]
  names <- turnedon$Code

  # Figure out the groups that have multiple ages classes in each stage (or
  # cohort)
  multiple_ages <- turnedon[turnedon$NumAgeClassSize>1, c(1,4,10)]
  num_multi_age <- dim(multiple_ages)[1]

  ntimesteps <- length(unique(nums_data$time))

  # For each species with multiple ages, use the calc_Z function to get time
  # varying Z values and put that all into one dataframe
  Z.dataframe <- data.frame()
  #for(i in 1:num_multi_age) {
    #temp_nums <- nums_data[nums_data$species==multiple_ages$Name[i],]
    temp_nums <- nums_data[nums_data$species %in% multiple_ages$Name,]
    temp_Z <- calc_Z(yoy=yoy, nums=temp_nums, fgs=fgs, biolprm=biolprm, toutinc=runprm$toutinc)
    #Z.dataframe <- rbind(Z.dataframe, temp_Z)
    Z.dataframe <- temp_Z
  #}

  # HACK for now, since some Z values are negative, I am replacing them
  # with a random number
  randomZ <- runif(length(which(Z.dataframe$atoutput<0)))
  Z.dataframe[which(Z.dataframe$atoutput<0),c("atoutput")] <- randomZ

  # rename the "atoutput" column, since it is really just the mortality, and
  # we want to retain it for the next step of merging
  colnames(Z.dataframe)[6] <- c("Z")

  # merge together nums_data and the output Z.dataframe:
  new_nums <- merge(nums_data, Z.dataframe[,c("species", "time", "Z")],
                    by=c("species", "time"), all.x=TRUE)

  # since we have to expand the number of rows for groups with multiple true
  # age classes, we will loop through species and time, creating a new list
  # that each element will be a single species in a single time step but
  # across ages and boxes (these wil all be put together in the end)
  temp.list <- list()

  #SKG subset turnedon and only loop through species present in the nums input
  turnedon_sub <- turnedon %>%
    filter(Name %in% new_nums$species)
  names <- turnedon_sub$Code

  for(i in 1:length(names)) {
    # looping all species -- those with or without multiple true ages

    group.i <- turnedon_sub$Name[i]
    nums_species <- new_nums[new_nums$species==group.i,] # might need to
    num_ages <- turnedon_sub$NumAgeClassSize[i]
    sp_times <- sort(unique(nums_species$time))
    n_sp_tsteps <- length(sp_times)
    # these last pieces are needed because not all species are present at all times

    # Check if multiple true ages, if not then just save species_nums to list
    if(num_ages==1) { temp.list[[length(temp.list)+1]] <- nums_species
    } else if(num_ages>1) {
      # take out the part of the Z.dataframe for species group i
      Zvals <- Z.dataframe[Z.dataframe$species==group.i,]

      # create empty list for this species i, each element in the list will be
      # a different time step
      list_species <- list()

      for(j in sp_times) { # looping through time
        # get the Z val for the timestep in question
        Zval.j <- Zvals$Z[Zvals$time==j]

        # and turn the Z value into a vector of survival values across the
        # number of true ages for each age class for this species
        nums_vec <- 1
        for(k in 1:(num_ages-1)) {
          nums_vec <- c(nums_vec, exp(-Zval.j*k))
        }
        nums_proportion <- nums_vec/sum(nums_vec)

        # take out the species numbers only for time step j
        nums_sp_subset <- nums_species[nums_species$time==j, ]

        # create an empty list, each element of this list will be a row from
        # the nums_sp_subset data frame that is split into multiple pieces
        # to make the atoutput column have the correct dimensions
        list_ages <- list()

        # loop through all the rows in nums_sp_subset
        for(l in 1:nrow(nums_sp_subset)) {
          nums_row <- nums_sp_subset[l,]
          # create new data frame with same columns, but just more rows to
          # account for all the true ages
          new_rows <- data.frame(species=nums_row$species,
                                 time=nums_row$time,
                                 agecl=seq((nums_row$agecl*num_ages - num_ages+1),
                                           nums_row$agecl*num_ages),
                                 polygon=nums_row$polygon,
                                 layer=nums_row$layer,
                                 atoutput=nums_row$atoutput*nums_proportion,
                                 Z=nums_row$Z
                                 )

          ### NOTE: the agecl piece is funny -- because now we are making many
          # more age classes for some... I think this works? But might want
          # another set of eyes to check my logic

          list_ages[[length(list_ages)+1]] <- new_rows
        }

        # now combine all the elements of the list for different rows and make
        # it an element of the list_species (to represent one time step)
        list_species[[length(list_species)+1]] <- do.call("rbind", list_ages)
      }
      temp.list[[length(temp.list)+1]] <- do.call("rbind", list_species)
    }
  }

  # Now just create the output
  finalout_withZ <- do.call("rbind", temp.list)

  ### CHECK:
  # do we want to remove the column of Z values? I think so
  finalout <- finalout_withZ[,-7]

  #return(finalout)

#}

That was me introducing a bug by filtering incorrectly; fixed it.

@@

# test with one species--this is still full polygon and layer output
trueNatage_ss <- trueNatage[trueNatage$species == species_ss,]

@@

len <- length_census_ss$natlength %>%
  filter(species == species_ss)

@@

truecatchbio_ss <- truecatchbio[truecatchbio$species == species_ss,]

@@

Now (that catch nums are corrected!--sort of; [when converted to catch in weight they are off by a factor of 2 for sardines in this model](https://sgaichas.github.io/poseidon-dev/SardinesOnlyatlantisom2SStest.html)) we can sample fishery catch for lengths, ages, weights, and as long as these are compositions and not absolute values they should be correct:

@@

Now we know the fishery output timestep, and there will only be 100 years. Because we don't want to wait 24 hours for this, we will look at only our three species.

@@

This is true catch at age for bocaccio (in age classes, not true ages):

@@

Can we split ages for the fishery comps too? True catch at age bocaccio:

```{r cattrueage}

YOY <- load_yoy(d.name, paste0("output", scenario.name, "YOY.txt"))

biol <- load_biolprm(d.name, biol.prm.file)

truecatage_ss <- calc_stage2age(nums_data = catch_testNss,
                                 biolprm = biol, 
                                 yoy = YOY,
                                 fgs = funct.groups, 
                                 runprm = runpar)

# run sample fish on this for completeness? yes to aggregate
truecatagesamp_ss <- sample_fish(truecatage_ss, effNhigh)

catchtrueage <- truecatagesamp_ss %>%
 filter(species == "Bocaccio_rockfish")

# plot it

catageplot <- ggplot(catchtrueage, aes(x=agecl, y=atoutput)) +
  geom_point() +
  theme_tufte() +
  labs(subtitle = paste(scenario.name,
                        catchtrueage$species))
